A Motion Capture and Imitation Learning-based Approach to Robot Control

### 2022-06-28 tex-arxiv/article.tex
add overview schematic, adjust image spacing in results

@@

Simulation allows recording exact pose information directly at regular time intervals, but setting up the scene necessitates creating a virtual environment where the task can be performed. In addition, for realistic interactions a physics simulation and immersive interface such as VR is required, but even this does not present the human demonstrator with the haptic feedback of actually performing the task in the real world. As such we decided that the benefits of simple setup and inherently natural interaction with the physical scene outweigh the precision and cost advantages of simulated environments, and selected motion capture as the source of demonstration data. For extracting additional information specific to the throwing task, we also track the object thrown and use its pose information to annotate each observation with target coordinates extrapolated from its flight arc and a release timing signal determined by heuristics described below.

@@

After an initial hyperparameter discovery process, the values in table \ref{tab:table1} were arrived at for both model types respectively. A range of model sizes and training epoch counts was compared for both models. In the case of the recurrent neural network, performance with different learning rates was also evaluated. As the feedforward models were developed first, performance with and without a time signal in the input data was also compared. For the RNN architecture, GRU was selected as prior research suggests that it outperforms LSTM when dealing with small data sets of long sequences \citep{yang2020lstm}.

\begin{table}
	\caption{Model hyperparameters}
	\centering
	\begin{tabular}{lll}
		\toprule
		Parameter & Feedforward & RNN \\
		\midrule
		Architecture & 2 fully-connected hidden layers, ReLU  & GRU, fully connected linear output   \\
		Parameter counts     & 128-1024 perceptrons per layer & 128-512 perceptrons in the unit \\
		Training epochs     & 20-100       & 300-1200  \\
		Batch size    & 64       & 32  \\
		Optimizer    & \multicolumn{2}{c}{Adam}  \\
		Learning rate    & $10^{-4}$       & $10^{-3}, 10^{-4}$   \\
		\bottomrule
	\end{tabular}
	\label{tab:table1}
\end{table}

@@

\begin{figure}[!b]
	\centering
	\fbox{\includegraphics[width=16cm]{../img/modeltype_comparison.png}}
	\caption{Results -- comparison between model architectures (\emph{Naive} -- refers to the feedforward architecture, also ``naive behavioral cloning''), the best performance attained in each class. Top row metrics -- cosine similarity, Euclidean distance (global); mean position error (stepwise). Bottom row -- mean rotation error (stepwise), mean throw error.}
	\label{fig:res_model_compare}
\end{figure}

\begin{figure}[!hbt]
	\centering
	\fbox{\includegraphics[width=16cm]{../img/each_model_best_perf.png}}
	\caption{Results -- best performance for each hyperparameter value. The top row shows feedforward model results, the bottom -- RNN. In both cases, the smaller, newer dataset with less variance in orientation and target coordinates results in better throw evaluations.  Introducing the time signal significantly improves feedforward model performance, whereas for the RNN improved results could be attained at higher learning rates. In the case of the feedforward model, training for too long and making the model too large initially appears to be detrimental, while for the RNN no clear trend was observed in this respect.}
	\label{fig:rnn_naive_best}
\end{figure}

@@

Fig. \ref{fig:vis_exec} illustrates how model outputs were visualized to use their qualitative aspects in guiding hyperparameter choice, and a preliminary implementation on a real robot was achieved -- executing trajectories generated by the simpler feedforward network, with target coordinates sampled from normal distributions corresponding to their values in the training data set. As can be seen, when these outputs were used to generate motion plans for a \emph{Universal Robotics UR5e} robot equipped with a pneumatic gripper, a plastic bottle was successfully thrown into a target container located outside the robot's reachable volume. Adequate output sequences were attained with both model architectures at multiple hyperparameter combinations, so to draw specific conclusions the quantitative evaluation metrics discussed in \ref{sec:eval} were produced for the 84 systematically trained models described above. Of the feedforward models, 18 out of 48 (37.5\%) successfully triggered a gripper release on every evaluation trajectory, enabling mean throw error estimates to be computed. Among the RNNs this was true for 24 out of 36 (66.7\%).

Figure \ref{fig:res_model_compare} shows a broad comparison between the RNN and feedforward model architectures, with each of the evaluation metric classes being present, though not all the specific metrics discussed -- as there was found to be a considerable amount of duplication in their findings. All figures shown correspond to the best result attained by either type in each of the performance indicators, irrespective of other variables. Figure \ref{fig:rnn_naive_best} elaborates upon each model class, showing the best attained results at various hyperparameter values. In the case of the feedforward model, the input observations did not initially contain a time signal, which was rectified early on. In the case of the RNN, deviating from the default learning rate was found to be beneficial. For both architectures, the best performance at each epoch count and parameter count is also plotted. Figure \ref{fig:ind_rnn_naive} shows the impact of only varying the crucial size (perceptron count) -- in each case the two best performing models at each fixed parameter combination are selected and their performance at different values of the independent variable is graphed.

>>>>

\begin{figure}
	\centering
	\fbox{\includegraphics[width=16cm]{../img/approach_schematic.png}}
	\caption{High level overview of the proposed approach. Demonstrations are recorded with motion capture equipment, a split and cropped data set augmented with target coordinates and a gripper actuation signal is created. This is used to train neural network models in behavioral cloning. When operating in the open-loop regime, prior model outputs are used to autoregressively predict subsequent states. The resulting sequence can then be fed into a Cartesian motion planner to control a robot.}
	\label{fig:approach}
\end{figure}

@@

Simulation allows recording exact pose information directly at regular time intervals, but setting up the scene necessitates creating a virtual environment where the task can be performed. In addition, for realistic interactions a physics simulation and immersive interface such as VR is required, but even this does not present the human demonstrator with the haptic feedback of actually performing the task in the real world. As such we decided that the benefits of simple setup and inherently natural interaction with the physical scene outweigh the precision and cost advantages of simulated environments, and selected motion capture as the source of demonstration data. For extracting additional information specific to the throwing task, we also track the object thrown and use its pose information to annotate each observation with target coordinates extrapolated from its flight arc and a release timing signal determined by heuristics described in the section on preprocessing \ref{sec:preprocess}.

@@

After an initial hyperparameter discovery process, the values in table \ref{tab:table1} were arrived at for both model types respectively. A range of model sizes and training epoch counts was compared for both models. In the case of the recurrent neural network, performance with different learning rates was also evaluated. As the feedforward models were developed first, performance with and without a time signal in the input data was also compared. For the RNN architecture, a \emph{gated recurrent unit} (GRU) was selected as prior research suggests that it outperforms \emph{long short-term memory} (LSTM) when dealing with small data sets of long sequences \citep{yang2020lstm}.

\begin{table}
	\caption{Model hyperparameters}
	\centering
	\begin{tabular}{lll}
		\toprule
		Parameter & Feedforward & RNN \\
		\midrule
		Architecture & 2 dense hidden layers, ReLU  & GRU, dense linear output   \\
		Parameter counts     & 128-1024 perceptrons per layer & 128-512 perceptrons in the unit \\
		Training epochs     & 20-100       & 300-1200  \\
		Batch size    & 64       & 32  \\
		Optimizer    & \multicolumn{2}{c}{Adam}  \\
		Learning rate    & $10^{-4}$       & $10^{-3}, 10^{-4}$   \\
		\bottomrule
	\end{tabular}
	\label{tab:table1}
\end{table}

@@

\newcommand{\reswidth}{15.95cm}

\begin{figure}[!b]
	\centering
	\fbox{\includegraphics[width=\reswidth]{../img/modeltype_comparison.png}}
	\caption{Results -- comparison between model architectures (\emph{Naive} -- refers to the feedforward architecture, also ``naive behavioral cloning''), the best performance attained in each class. Top row metrics -- cosine similarity, Euclidean distance (global); mean position error (stepwise). Bottom row -- mean rotation error (stepwise), mean throw error.}
	\label{fig:res_model_compare}
\end{figure}

\begin{figure}[!hbt]
	\centering
	\fbox{\includegraphics[width=\reswidth]{../img/each_model_best_perf.png}}
	\caption{Results -- best performance for each hyperparameter value. The top row shows feedforward model results, the bottom -- RNN. In both cases, the smaller, newer dataset with less variance in orientation and target coordinates results in better throw evaluations.  Introducing the time signal significantly improves feedforward model performance, whereas for the RNN improved results could be attained at higher learning rates. In the case of the feedforward model, training for too long and making the model too large initially appears to be detrimental, while for the RNN no clear trend was observed in this respect.}
	\label{fig:rnn_naive_best}
\end{figure}

@@

Fig. \ref{fig:vis_exec} illustrates how model outputs were visualized to use their qualitative aspects in guiding hyperparameter choice, and a preliminary implementation on a real robot was achieved -- executing trajectories generated by the simpler feedforward network, with target coordinates sampled from normal distributions corresponding to their values in the training data set. As can be seen, when these outputs were used to generate motion plans for a \emph{Universal Robots UR5e} robot equipped with a pneumatic gripper, a plastic bottle was successfully thrown into a target container located outside the robot's reachable volume. Adequate output sequences were attained with both model architectures at multiple hyperparameter combinations, so to draw specific conclusions the quantitative evaluation metrics discussed in \ref{sec:eval} were produced for the 84 systematically trained models described above. Of the feedforward models, 18 out of 48 (37.5\%) successfully triggered a gripper release on every evaluation trajectory, enabling mean throw error estimates to be computed. Among the RNNs this was true for 24 out of 36 (66.7\%).

Figure \ref{fig:res_model_compare} shows a broad comparison between the RNN and feedforward model architectures, with each of the evaluation metric classes being present, though not all the specific metrics discussed -- as there was found to be a considerable amount of duplication in their findings. All figures shown correspond to the best result attained by either type in each of the performance indicators, irrespective of other variables.

Figure \ref{fig:rnn_naive_best} elaborates upon each model class, showing the best attained results at various hyperparameter values. In the case of the feedforward model, the input observations did not initially contain a time signal, which was rectified early on. In the case of the RNN, deviating from the default learning rate was found to be beneficial. For both architectures, the best performance at each epoch count and parameter count is also plotted. Figure \ref{fig:ind_rnn_naive} shows the impact of only varying the crucial size (perceptron count) -- in each case the two best performing models at each fixed parameter combination are selected and their performance at different values of the independent variable is graphed.
